---
title: "Lab4"
output:
  html_document: default
  pdf_document: default
date: "2022-12-21"
editor_options:
  markdown:
    wrap: 72
---

Team:
- Bilinskyi Andrii
- Markiyan Kruhliy
- Mokrytskyi Liubomyr


```{r}
require(BSDA)
library(BSDA)
require(EnvStats)   
library(EnvStats)
```

The data for problems 1--3 are generated as follows: set ak := {k ln
(k2n + π)}, k ≥ 1

xk = Φ−1 (ak), k = 1, . . . , 100, yl = Φ−1 (al+100), l = 1, . . . , 50

```{r}
id <- 11
set.seed(id)

a.data <- c()


a_k_calculate <- function(k) {
  a <- k * log(k^2 * id + pi)
  return(a - trunc(a))
}


a.data <- a_k_calculate(1:150)


x.data <- qnorm(a.data[1:100])
y.data <- qnorm(a.data[101:150])

```

```{r}
a_k_calculate(1)
```

PROBLEM 1

In this case we need to test H0 : µ1 = µ2 vs. H1 : µ1 ̸= µ2; σ1 = 1; σ2
= 1

To test mean of two normal distributions with known variance, we use
Two-sided z-test.

Lets estimate µ1 - µ2:

$C_\alpha= \{(\mathbf x, \mathbf y) : |{\mathbf X}-\overline {\mathbf Y}| \gt c_\alpha \}$

$Z = Z(\mathbf X,\mathbf Y) := \sqrt{\frac{mn}{(m+n)}} \frac{\overline{\mathbf x} - \overline{\mathbf y}}{\sigma}$

$\boxed{\strut\ |z(\mathbf x,\mathbf y)|\ge z_{1-\alpha/2} \ }$

for ${\alpha}=0.05$:

$|z(\mathbf x,\mathbf y)|\ge z_{0.975}$

$H_0$ should be rejected on the significance level 0.05 but p-value of
the test is 0.1943 so $H_0$ should not be rejected for sure.

```{r}
z.test(x = x.data, y = y.data, alternative = "two.sided", sigma.x = 1, sigma.y = 1)
```

Problem 2

Hypothesis:
$H_0\,: \sigma^2_x = \sigma_y^2 \quad \textrm{vs} \quad H_1\,: \sigma^2_x > \sigma_y^2$
Here we perform an f tests to compare the variances of two samples from
populations, by dividing them.

The rejection region is:
$C_\alpha := \{ \mathbf x\in\mathbb R^n,\mathbf y\in\mathbb R^m \mid f(\mathbf x,\mathbf y)\le f_{\alpha} \}$
where $f_{\alpha}$ is the quantile of level ${\alpha}$ for the Fischer
distribution $\mathscr{F}_{n-1,m-1}$ and ${\alpha}=0.05$.

p-value of is equals 0.7434 so $H_0$ should not be rejected.

```{r}
var.test(x = x.data, y = y.data, alternative = "g");
```

PROBLEM 3

$H_0$: $\{x_k\}^{100}_1$ are normally distributed vs. $H_1$:
$\{x_k\}^{100}_1$ aren't normally distributed To test hypothetical
distribution we should use Kolmogorov's Goodness-of-fit test. By LLN if
sample cdf $\hat{F}_x$ is close to cdf of normal distribution then our
statistics $d = \sup_{t\in R}|\hat{F}_x(t)-Ф(t)|$ should take small
values. Here rejection region
$C_\alpha = \{x\in R| d\ge d_{1-\alpha}^{(n)}\}$

As p-value usually equals to 0.1-0.9 (it changes because normal
distributed sample changes), our first type error is small enough, so we
should accept $H_0$ hypothesis.

```{r}
var.test(x.data, y.data, alternative = "g")
ks.test(x.data, "pnorm", mean = mean(x.data), sd = sd(x.data))
ks.test(abs(x.data), "pexp", 1)
ks.test(x.data, y.data)
```
